---
title: "Intro to R"
author: "Marie Mortensen"
date: "1/2/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1) 
Use R to figure out how many elements in the vector below are greater than 2 . (You need to filter out the NAs first)

```{r}
rooms <- c(1, 2, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, NA)
#filtering out NAs
rooms <- rooms[!is.na(rooms)]

#using indexing ([]) and length() to count the number of room elements above 2
length(rooms[rooms>2])

```
There are 8 elements  above 2

## 2) 
What is the result of running median() function on the above 'rooms' vector? (again, best remove the NAs)

```{r}
pacman::p_load(tidyverse)
rooms %>% median() 
```
Median is 1.5

## 3) 
Inside your R Project (.Rproj), install the 'tidyverse' package and use the download.file() and read_csv() function
to read the SAFI_clean.csv dataset into your R project as 'interviews' digital object
(see instructions in https://datacarpentry.org/r-socialsci/setup.html and 'Starting with Data' section).
Take a screenshot of your RStudio interface showing
a) the script you used to create the object,
b) the 'interviews' object in the Environment and the
c) structure of your R project in the bottom right Files pane.

```{r}
pacman::p_load(tidyverse)
download.file("https://ndownloader.figshare.com/files/11492171",
              "../data/SAFI_clean.csv", mode = "wb")
interviews <- read.csv("../data/SAFI_clean.csv", na = "NULL")
```


